---
title: "Analysis"
author: "Keisuke Atsumi"
date: "`r format(Sys.time(), '%Y %m %d')`"
output:
  html_document:
    code_folding: hide
    theme: united
    toc: yes
    toc_float: yes
---

# Setup


Load packages & original functions

- `summarise_simul` : Load & output simulation result
- `make_config` : Make simulation config file

```{r}

rm(list=ls())  # reset workspace

pacman::p_load(
 tidyverse
 , RcppRoll  # window functions
 , reshape2
 , foreach   # Pararel
 , knitr     # rmarkdown
 # , magrittr  # Extend pipe
 , kableExtra # nice tables
 , pander     # nice tables
 # , patchwork
 )

opts_chunk$set(prompt=TRUE, message=FALSE, comment="", warning = FALSE) 
options(knitr.kable.NA = '', scipen=100)  # do not show numbers using exponential

# Make inherit style table
inherit_table <- tibble(
  "rowname" = c("TT", "Tt", "tt"),
  "PP" = NA, "Pp" = NA, "pp" = NA
  ) %>% 
  column_to_rownames("rowname")


#' Title: ext_count
#' Load simulation result and Extract individual count data for each phenotype
#' @param gset: Genome setting
#' @param sexsel: Sexual selection style
#' @param gen: Simulated generations
ext_count <- function(gset = "", sexsel = "", gens = c(1:100)) {
  
  for(bdmi in c(0, 0.1, 0.4, 0.8)) {
    for(ud in c(0, 0.1, 0.4, 0.8)) {
      
      # Load simulation data. Repeat for all generation files
      foreach(iter = c("0","0_1","0_10","0_100","0_11","0_12","0_13","0_14","0_15","0_16","0_17","0_18","0_19","0_2","0_20","0_21","0_22","0_23","0_24","0_25","0_26","0_27","0_28","0_29","0_3","0_30","0_31","0_32","0_33","0_34","0_35","0_36","0_37","0_38","0_39","0_4","0_40","0_41","0_42","0_43","0_44","0_45","0_46","0_47","0_48","0_49","0_5","0_50","0_51","0_52","0_53","0_54","0_55","0_56","0_57","0_58","0_59","0_6","0_60","0_61","0_62","0_63","0_64","0_65","0_66","0_67","0_68","0_69","0_7","0_70","0_71","0_72","0_73","0_74","0_75","0_76","0_77","0_78","0_79","0_8","0_80","0_81","0_82","0_83","0_84","0_85","0_86","0_87","0_88","0_89","0_9","0_90","0_91","0_92","0_93","0_94","0_95","0_96","0_97","0_98","0_99"), .combine = 'rbind') %do% {
        
        #Because generation length vary among params & iters, obtain generation length 
        maxgen <- length(list.files(
          paste0("D:/Simulation/Result/", gset, "_", sexsel, "_b", bdmi, "h", ud, "/", iter)
            ))/2-1
        
        foreach(g = 1:maxgen, .combine = 'rbind') %dopar% {  
          # Read tab delimited files for phenotype
          read_tsv(paste0("D:/Simulation/Result/", gset, "_", sexsel, "_b", bdmi, "h", ud, "/", iter, "/Gen", g, "_phenotypes.txt"), show_col_types = FALSE) %>%
            as_tibble %>%  
            filter(pop == 3) %>%                     # Use only hybrid population
            dplyr::select("Signal0") %>% 
            mutate_at("Signal0", as.numeric) %>%     # Group by phenotype
            mutate_all(as.factor) %>%
            group_by(Signal0) %>% 
            summarise(nPheno = length(Signal0)) %>%  # Count indv for each phenotype
            mutate(
              nGen = sum(nPheno),
              Generation = g
              ) 
          } %>% 
          mutate(Iter = iter) #Add column indicating iteration
        } %>% 
        mutate(Freq = nPheno/nGen*100) %>% 
        write_csv(paste0("Data/count_", gset, "_", sexsel, "_b", bdmi, "h", ud, ".csv"))
    }}
  
  }


#' Title: ext_stat
#' Load simulation result and Extract mean & variance from phenostats file
#' @param gset: Genome setting
#' @param sexsel: Sexual selection style
#' @param gen: Simulated generations
ext_stat <- function(gset = "", sexsel = "") {  #, gens = c(1:100)
  
  for(bdmi in c(0, 0.1, 0.4, 0.8)) { 
    for(ud in c(0, 0.1, 0.4, 0.8)) {
      
      # Load simulation data. Repeat for all generation files
      foreach(iter = c("0","0_1","0_10","0_100","0_11","0_12","0_13","0_14","0_15","0_16","0_17","0_18","0_19","0_2","0_20","0_21","0_22","0_23","0_24","0_25","0_26","0_27","0_28","0_29","0_3","0_30","0_31","0_32","0_33","0_34","0_35","0_36","0_37","0_38","0_39","0_4","0_40","0_41","0_42","0_43","0_44","0_45","0_46","0_47","0_48","0_49","0_5","0_50","0_51","0_52","0_53","0_54","0_55","0_56","0_57","0_58","0_59","0_6","0_60","0_61","0_62","0_63","0_64","0_65","0_66","0_67","0_68","0_69","0_7","0_70","0_71","0_72","0_73","0_74","0_75","0_76","0_77","0_78","0_79","0_8","0_80","0_81","0_82","0_83","0_84","0_85","0_86","0_87","0_88","0_89","0_9","0_90","0_91","0_92","0_93","0_94","0_95","0_96","0_97","0_98","0_99"), .combine = 'rbind') %do% {
        
        #Because generation length vary among params & iterations, obtain generation length 
        maxgen <- length(list.files(
          paste0("D:/Simulation/Result/",gset,"_",sexsel,"_b",bdmi,"h",ud,"/", iter)
            ))/2-1
        
        foreach(g = 1:maxgen, .combine = 'rbind') %dopar% {  
        
          # Read tab delimited files for phenotype from HDD
          read_tsv(paste0("D:/Simulation/Result/", gset, "_", sexsel, "_b", bdmi, "h", ud, "/", iter, "/Gen", g, "_phenostats.txt"), show_col_types = FALSE, col_names = FALSE) %>%
            as_tibble %>%  
            filter(
              .[1] == 3,
              .[2] == "Hybridindex" | .[2] == "Pref0" | .[2] == "Signal0"
              ) %>% 
            mutate(Generation = g)
          } %>% 
          mutate(Iter = iter) #Add column indicating iteration
        } %>% 
        select(-"X1") %>%         
        write_csv(paste0("Data/mean_", gset, "_", sexsel, "_b", bdmi, "h", ud, ".csv"))
  
    }}
  
  }


# Plot settings
mytheme <- theme_bw() + # simple theme
  theme(panel.grid.major.x = element_blank(),
        axis.title = element_text(size=12),
        axis.text = element_text(size=11),
        axis.text.x = element_text(angle = 30, hjust = 1),
        strip.text = element_text(size=12, lineheight=5.0),
        strip.background = element_rect(fill = "lightgrey")
        ) 

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#0072B2", "#D55E00", "#CC79A7")

```



```{r}

ext_stat(gset = "dis_unlink", sexsel = "disruptive")



```





## Result

- 2 loci models : Preference and traits are determined by each 1 gene on different chromosome

- 1 locus models : Preference and traits are determined by 1 gene

### Disruptive sexual selection

```{r}

# Visualize inheritance pattern
inherit_table$pp <- c(0.245,0.01,0.745)
inherit_table$Pp <- c(0.495,0.01,0.495)
inherit_table$PP <- c(0.745,0.01,0.245)
inherit_table %>% 
  kable("html", caption = "**Trait \\ Preference**") %>% 
  kable_styling("striped", position = "left", full_width = F)

inherit_table %>% 
  rownames_to_column("Trait") %>% 
  pivot_longer(!Trait, names_to = "Preference", values_to = "Prob") %>% 
  within(Preference <- factor(Preference, levels = c("PP", "Pp", "pp"))) %>% 
  within(Trait <- factor(Trait, levels = c("tt", "Tt", "TT"))) %>% 
  ggplot(aes(Preference, Trait, fill= Prob)) + 
    geom_tile(colour = "grey80") +
    scale_fill_gradient(low = "white", high = "#E60012") +
    scale_x_discrete(position = "top") +
    theme_void() +
    theme(axis.text = element_text(color = "black"))
  

cap = "Disatractive hybrids"
gset = "dis_link"
sexsel = "flip"

data <- foreach(bdmi = c(0, 0.1, 0.4, 0.8), .combine = 'rbind') %do% {
  foreach(ud = c(0, 0.1, 0.4, 0.8), .combine = 'rbind') %do% {
    tmp <- read_csv(show_col_types=FALSE,
      paste0("Data/mean_", gset, "_", sexsel, "_b", bdmi, "h", ud, ".csv")) %>% 
      filter(X2 == "Signal0") %>% 
      select("Iter", "Generation", "X3") %>% 
      mutate_at("Iter", as.factor) %>% 
      group_by(Iter)
    tmp %>% 
      filter(X3 == "1[0]" | X3 == "0[0]") %>%     #Generations with fixed genotype
      filter(Generation == min(Generation)) %>%   #Time that genotype was fixed
      rename("Gen_fix" = "Generation", "Genotype" = "X3") %>%  #
      left_join(summarise(tmp, Gen_ext = n()), ., by = "Iter") %>%  #Combine with simul length
      mutate(s_underdominance = as.factor(ud))  
    } %>% 
    mutate(s_bdmi = as.factor(bdmi))  
  } %>% 
  mutate(s = str_c(s_underdominance, s_bdmi, sep = ", "))

ggplot(data, aes(x = s, y = Gen_fix)) +
  geom_violin(fill = "grey80", colour = NA) +
  geom_point(aes(colour = Genotype), position = "jitter", size = 1) +
  xlab("s (Underdominance, BDMI)") +  ylab ("Generations to fix") +
  mytheme

```

<div style="margin-bottom:40px;">
</div>


### Self-reference assortative matings 

```{r}
# Visualize inheritance pattern
inherit_table$pp <- c(0.005,0.005,0.99)
inherit_table$Pp <- c(0.005,0.99,0.005)
inherit_table$PP <- c(0.99,0.005,0.005)
# Output inheritance style table
inherit_table %>% 
  kable("html", caption = "**Trait \\ Preference**") %>% 
  # here you can add the vertical line
  column_spec (2:4, border_left = T, border_right = T) %>%
  kable_styling(position = "left", full_width = F)

cap = "Self-reference assortative mating"


dat %>% ggplot(aes(x = Generation, y = Freq, colour = Signal0, group = Signal0)) +
  geom_line() +
  labs(y = "Frequency (%)", colour = "Courter phenotype", title = cap) +
  facet_grid(Naturalselection ~ Locus, scales = "free")

```

<div style="margin-bottom:40px;">
</div>

### Recessive preference

```{r}

# Visualize inheritance pattern
inherit_table$PP <- c(0.99,0.005,0.005)
inherit_table$Pp <- c(0.005,0.005,0.99)
inherit_table$pp <- c(0.005,0.005,0.99)
inherit_table %>% 
  kable("html", caption = "**Trait \\ Preference**") %>% 
  # here you can add the vertical line
  column_spec (2:4, border_left = T, border_right = T) %>%
  kable_styling(position = "left", full_width = F)
cap = "Recessive preference"

dat <- foreach(locus = 1:2, .combine = 'rbind') %do% {
  foreach(ns = c(0.3, 1), .combine = 'rbind') %do% {
    load_simul(l = locus, sexsel = "recessive", natsel = ns)
    } } %>% 
  summarise_simul()
levels(dat$Locus) <- c("Linked", "Unlinked")

dat %>% ggplot(aes(x = Generation, y = Freq, colour = Signal0, group = Signal0)) +
  geom_line() +
  labs(y = "Frequency (%)", colour = "Courter phenotype", title = cap) +
  facet_grid(Naturalselection ~ Locus, scales = "free")

```

<div style="margin-bottom:40px;">
</div>

### Pseudo-dominance

```{r}

# Visualize inheritance pattern
inherit_table$PP <- c(0.99,0.005,0.005)
inherit_table$Pp <- c(0.99,0.005,0.005)
inherit_table$pp <- c(1/3,1/3,1/3)
inherit_table %>% 
  kable("html", caption = "**Trait \\ Preference**") %>% 
  kable_styling("striped", position = "left", full_width = F)

cap = "Pseudo-dominance"

dat <- foreach(locus = 1:2, .combine = 'rbind') %do% {
  foreach(ns = c(0.3, 1), .combine = 'rbind') %do% {
    load_simul(l = locus, sexsel = "pdominance", natsel = ns)
    } } %>% 
  summarise_simul()
levels(dat$Locus) <- c("Linked", "Unlinked")

dat %>% ggplot(aes(x = Generation, y = Freq, colour = Signal0, group = Signal0)) +
  geom_line() +
  labs(y = "Frequency (%)", colour = "Courter phenotype", title = cap) +
  facet_grid(Naturalselection ~ Locus, scales = "free")

```

Heterozygotes and recessive are favored (why??)

<div style="margin-bottom:40px;">
</div>

### Heterozygotes mate randomly

```{r}

# Visualize inheritance pattern
inherit_table$PP <- c(0.99,0.005,0.005)
inherit_table$Pp <- c(1/3,1/3,1/3)
inherit_table$pp <- c(0.005,0.005,0.99)
inherit_table %>% 
  kable("html", caption = "**Trait \\ Preference**") %>% 
  kable_styling("striped", position = "left", full_width = F)

cap = "Heterozygotes mate randomly"

dat <- foreach(locus = 1:2, .combine = 'rbind') %do% {
  foreach(ns = c(0.3, 1), .combine = 'rbind') %do% {
    load_simul(l = locus, sexsel = "heterorandom", natsel = ns)
    } } %>% 
  summarise_simul()
levels(dat$Locus) <- c("Linked", "Unlinked")

dat %>% ggplot(aes(x = Generation, y = Freq, colour = Signal0, group = Signal0)) +
  geom_line() +
  labs(y = "Frequency (%)", colour = "Courter phenotype", title = cap) +
  facet_grid(Naturalselection ~ Locus, scales = "free")

```

Recessive is favored, and heterozygote is not favored

<div style="margin-bottom:40px;">
</div>

### Flip preference

```{r}

# Visualize inheritance pattern
inherit_table$PP <- c(0.99,0.005,0.005)
inherit_table$Pp <- c(0.49,0.01,0.49)
inherit_table$pp <- c(0.005,0.005,0.99)
inherit_table %>% 
  kable("html", caption = "**Trait \\ Preference**") %>% 
  kable_styling("striped", position = "left", full_width = F)

cap = "Flip preference"

dat <- foreach(locus = 1:2, .combine = 'rbind') %do% {
  foreach(ns = c(0.3, 1), .combine = 'rbind') %do% {
    load_simul(l = locus, sexsel = "flip", natsel = ns)
    } } %>% 
  summarise_simul()
levels(dat$Locus) <- c("Linked", "Unlinked")

dat %>% ggplot(aes(x = Generation, y = Freq, colour = Signal0, group = Signal0)) +
  geom_line() +
  labs(y = "Frequency (%)", colour = "Courter phenotype", title = cap) +
  facet_grid(Naturalselection ~ Locus, scales = "free")

```

<div style="margin-bottom:40px;">
</div>

### Atractive hybrids

```{r}

# Visualize inheritance pattern
inherit_table$PP <- c(0.245,0.745,0.01)
inherit_table$Pp <- c(0.005,0.99,0.005)
inherit_table$pp <- c(0.01,0.745,0.245)
inherit_table %>% 
  kable("html", caption = "**Trait \\ Preference**") %>% 
  kable_styling("striped", position = "left", full_width = F)

cap = "Atractive hybrids"

dat <- foreach(locus = 1:2, .combine = 'rbind') %do% {
  foreach(ns = c(0.3, 1), .combine = 'rbind') %do% {
    load_simul(l = locus, sexsel = "speriorhyb", natsel = ns)
    } } %>% 
  summarise_simul()
levels(dat$Locus) <- c("Linked", "Unlinked")

dat %>% ggplot(aes(x = Generation, y = Freq, colour = Signal0, group = Signal0)) +
  geom_line() +
  labs(y = "Frequency (%)", colour = "Courter phenotype", title = cap) +
  facet_grid(Naturalselection ~ Locus, scales = "free")

```

<div style="margin-bottom:40px;">
</div>

### Oblique positively learned preference

```{r}

cap = "Oblique positive"

dat <- foreach(ns = c(0.3, 1), .combine = 'rbind') %do% {
  load_simul(l = 8, sexsel = "obl_posit", natsel = ns)
  } %>% 
  summarise_simul()

dat %>% ggplot(aes(x = Generation, y = Signal0, size = Freq)) +
  geom_point() +
  labs(y = "Frequency (%)", title = cap) +
  facet_grid(Naturalselection ~ Locus, scales = "free")

```

<div style="margin-bottom:40px;">
</div>

### Oblique negatively learned preference

```{r}

cap = "Oblique negative"

dat <- foreach(ns = c(0.3, 1), .combine = 'rbind') %do% {
  load_simul(l = 8, sexsel = "obl_negat", natsel = ns)
    } %>% 
  summarise_simul()

dat %>% ggplot(aes(x = Generation, y = Signal0, size = Freq)) +
  geom_point() +
  labs(y = "Frequency (%)", title = cap) +
  facet_grid(Naturalselection ~ Locus, scales = "free")

```


```{r}

# ggplot(aes(x = Generation, y = Freq, colour = Signal0, group = Signal0)) +
#   geom_line() +
#   labs(y = "Frequency (%)", colour = "Courter phenotype")

  # assign(
  #   paste0("plot_d_", locus),
  #   # Density graph
  #   ggplot(dat, aes(x = Signal0, colour = Generation)) +
  #     geom_density() +
  #     labs(
  #       x = "Courter phenotype", 
  #       title = paste(locus, "locus |", cap)
  #       )
  #   )

```


